### 2015-05-05 background.tex
more updates in the background section

\ADComment{Can we justify the focus on locks a bit more?  E.g., are you able to survey, via grepping, what percentage of Linux drivers make use of locks (likely 100\%?) vs.\ what percentage also make use of lock-free mechanisms?}

@@

In the example of Figure~\ref{fig:locksets}, we analyze only one thread $T$ and one shared variable $A$ and thus there are two locksets: the current lockset for $T$ ($\mathit{CLS}_T$) and the lockset for $A$ ($\mathit{LS}_A$).

Figure~\ref{fig:locksets} shows an example of applying lockset analysis on a concurrent program consisting of two threads $T_1$ and $T_2$, both accessing the shared variable $A$. Thus, we have to analyze three locksets: the current locksets $\mathit{CLS}_1$ and $\mathit{CLS}_2$ for $T_1$ and $T_2$ respectively; and the lockset $\mathit{LS}_A$ for $A$.

\ADComment{I also think you should use a two-threaded example, instead of a one-threaded example, to show that although thread $T$ might consistently use mutex $M$ to protect variable $A$, this is no good if thread $U$ accesses variable $A$ without holding $M$.}

>>>>

We believe that our focus in locks is justified, because the majority of Linux drivers still use locks. As a sanity check, we used a simple grep command and found that 68\% of the drivers from the 4.0 Linux kernel distribution use locks, versus 13\% that use atomic operations. We also found that 31\% of the drivers use neither locking nor atomic primitives, which suggests that they are either simple (non-concurrent) drivers or use some other lock-free synchronization primitive.

@@

Figure~\ref{fig:locksets} shows an example of applying lockset analysis on a concurrent program consisting of two threads $T_1$ and $T_2$, both accessing the same shared variable $A$. Thus, we have to track three locksets: the current locksets $\mathit{CLS}_1$ and $\mathit{CLS}_2$ for $T_1$ and $T_2$ respectively; and the lockset $\mathit{LS}_A$ for $A$. Initially, $\mathit{LS}_A$ contains the two locks $M$ and $N$ that the two threads acquire during execution. When $T_1$ terminates, $\mathit{LS}_A$ contains only the lock $N$, because the thread write accessed $A$ without holding $N$. Next, during the execution of $T_2$, the thread write accesses $A$ without holding $M$. Thus, $\mathit{LS}_A$ becomes the empty set. This means that there is a potential data race between the two threads on $A$ as $T_2$ does not consistently protect $A$ with at least one common with $T_1$ lock.
